\chapter{Analysis and Lessons Learned}
\index{Analysis and Lessons Learned%
@\emph{Analysis and Lessons Learned}}%
\label{ch:analysis}

This report aims to demonstrate how certain traditionally difficult problems in web development have been addressed by Web Components\index{Web Components}.
The principles underlying Web Components and Polymer\index{Polymer} are summarized by the list in~\cref{sec:wcprinciples} taken from ~\cite{webcomponentscontributors2014}.
I have tried to make Speakur\index{Speakur} adhere to these principles both internally and externally,
and to demonstrate the application of abstraction\index{abstraction} and encapsulation\index{encapsulation} patterns to creating flexible, composable, reusable, sharable components.
My goal is to understand how the W3C Web Component\index{Web Components} initiative will impact web application engineering\index{software engineering} in the future.
By all appearances, the long term impact will be substantial.

\section{Web Components Architecture Analysis}
Web Component's\index{Web Components} essential benefit is that it allows decoupling of components, 
separation of concerns, 
and the encapsulation of internal structure.
This raises bundles of discrete functionality or `components' to a higher level of abstraction and 
will help open up a new ecosystem of reusable, extendable, sharable modules.
The Web Component principles from~\cref{sec:wcprinciples} emphasize thinking small, 
breaking the problem down into manageable chunks and being adaptable, extendable and composable.
Web developers were already doing this in many cases, but without formal support from the browser.

Perhaps the most interesting piece of advice is to ``deliver the key benefit to HTML authors, not just coders''~\cite{webcomponentscontributors2014}.
It underscores that, fundamentally, 
the web consists of \textit{documents} with \textit{authors}, 
a fact that is sometimes forgotten by the myriad of JavaScript\index{JavaScript} libraries and frameworks\index{framework}.
The components that will find the most success will be those which focus first and foremost on providing convenience and benefit to HTML\index{HTML} authors and content developers, 
not just the software engineers\index{software engineering} who maintain the libraries and frameworks behind the scenes.
Speakur\index{Speakur} aims to provide as small a deployment surface as possible to web authors so they can easily add a discussion forum to their site without embedding a big blob of HTML.
The growth of social coding sites like GitHub\index{GitHub} will further encourage the sharing and reuse of Web Components designed along these principles.

\section{Writing Web Components with Polymer}
This section details some of the lessons learned developing Speakur\index{Speakur} on top of Google's\index{Google} Polymer\index{Polymer} framework.

\subsection{Using Data-Bound Templates}
Polymer's data bindings\index{data-bound template} are a powerful feature for driving `live' web applications.
Their utility extends well beyond user interface\index{user interface (UI)} views.
A 3-way automatic binding between the local model, the UI view, and the remote database is an extremely powerful abstraction,
even more so when combined with Firebase's\index{Firebase} 
Web\-Socket-based\index{WebSockets} event notification\index{event notification} system.
JavaScript variables can automatically update to reflect changes made by users on the other side of the world.
Of course, care must be taken to design the interface so the user doesn't accidentally make changes that are immediate and can't be undone. 
For example, Speakur's user preferences screen doesn't commit changes to the database until the user clicks the Save button.
Editing the text of an existing post also saves a copy of the previous version.

\subsection{Decoupling Components}
Data bindings and synchronization are discussed in~\cref{sec:sync}.
In general, two-way data bindings should only be used \textit{within} a component.
Across component boundaries, use a one-way binding to send data down to a child,
and fire events to notify containers of state changes.
DOM Mutation Observers\index{mutation observers} can be useful for making one component react to a change in another without directly tying their code together.
As the example with \tcode{lc} (locale) in~\cref{sec:i18n} showed, 
sometimes it is necessary to `shoehorn' certain variables into a Polymer binding expression 
to ensure that the expression is recalculated when that variable changes.

\subsection{Optimizing Polymer Performance}
Component load times are currently one problem area for Web Components.
The Vulcanize\index{Vulcanize} tool offers definite benefits for load times as seen in~\cref{table:speakurloading},
but Speakur is a relatively simple application where all 20 or so components are used more or less immediately.
The Vulcanize tool must be applied at the level of the site or app, not by the individual component authors.
Larger single page applications (SPAs) will face more complex challenges when trying to optimize load times, and performance in general,
before Web Components\index{Web Components} are fully native in browsers, especially when combining frameworks.

The \tcode{<template>}\index{<template>} element is a powerful new tool for web authors, 
but current implementations can have performance issues when rendering large lists of items, such as with a long list of comments. 
Polymer's\index{Polymer} \tcode{<core-list>} is more optimized for this use case.
Also, the \tcode{if=} attribute on nested \tcode{<template>}\index{<template>} elements should not be used in those cases where you could use the \tcode{hidden?=} attribute on the inner content instead.
There are several reasons for this, including performance and 
the fact that Polymer's handy \tcode{this.\$} \tcode{id}-to-object map is not dynamic; 
that is, element objects created conditionally in inner templates are not automatically added to \tcode{this.\$}.

\subsection{Inheritance in Polymer}
I recommend avoiding the use of Polymer's\index{Polymer} inheritance unless there is a clear case for providing polymorphism-like\index{polymorphism} behavior.
One reason for that is that (as of this writing) when a base class provides computed properties, 
these property definitions have to be repeated in any subclasses if any new ones are required,
which violates the ``don't repeat yourself'' (DRY)\index{don't repeat yourself (DRY)} principle.
Speakur uses Polymer inheritance\index{inheritance} in its CSS\index{CSS} theme components; 
most of the functionality is defined in the base class and the subclasses mainly provide numeric parameters for things like colors and margins.
Composition\index{composition} could achieve these behaviors almost as easily.

%Having multiple <core-overlay> active at the same time can cause problems. 
% I created a simple fix that adjusts the CSS \tcode{z-index} property. 
%You can find the fix here:~\cite{landers2015-d}.

\section{Lessons Learned}
When I started this project, I wasn't sure where it would take me. 
I was particularly unsure about the extent to which I could leverage the vast existing library of JavaScript frameworks and libraries when writing a new-style Web Component. Here are some of the lessons I learned while developing Speakur:
%\pagebreak
\begingroup
\setlist{parsep=0pt,listparindent=\parindent}
\begin{itemize}
\item \textbf{Web Components provide powerful abstractions.}

Using Web Components, and particularly the data-bound templates provided by Polymer, greatly reduced the amount of effort needed to code a responsive, live web app compared to my previous efforts, 
yet the result is still flexible and works in a variety of software architectures.
It's difficult to estimate that difference in terms of lines of code, but it is clear that Web Components offer better tools for code organization, 
separation of concerns, and modularization than were previously available.

\item \textbf{Web Components are standardized.}

Web Components follow standard HTML and HTTP\index{HTTP} semantics better than many previous JavaScript frameworks. 
A consistent component packaging and import system, the use of HTML attributes for API\index{API} specification, 
and the lack of a required compilation or preprocessing step all help Web Components fit cleanly into the architecture of the web.

\item \textbf{Use Firebase as an adjunct.}

Firebase\index{Firebase} is excellent for what it is, 
but for a really substantial or business-critical application it's probably best to treat it as a handy side-channel for data synchronization and event notification, 
rather than as the main business data repository. 
Aside from cost factors or the security issues inherent in storing your data in the cloud, 
running your own server remains the best way to ensure effective data integrity, control and validation.

\item \textbf{Polymer can have problems with older JavaScript frameworks.}

The ease of combining Polymer-based components with other frameworks and non-Web-Component-aware libraries (``legacy code'') depends heavily on what they do.
Using pure-CSS libraries (like \tcode{bootstrap.css}) with Polymer/WC is safe and easy to control by limiting their scope to a particular component. 

Writing a Polymer component that directly uses legacy JavaScript\index{JavaScript} libraries (like \tcode{bootstrap.js}) is not nearly as easy in many cases. 
I tried to use several Bootstrap\index{Bootstrap} JavaScript widgets but quickly ran into incompatibilities with treatment of the document model, event handling, and so on.
The Twitter Bootstrap developers, perhaps understandably, don't care that much about adding compatibility hacks for relatively immature projects like Polymer. 
Libraries that provide purely logical services like i18next\index{i18next}, instead of document or interface-related features like Bootstrap, are much easier to integrate.

\item \textbf{Include mobile early in the development cycle.}

For example, be careful where you apply margins when designing for mobile. 
If you have a group of interface widgets that form a logical unit, like a toolbar,
and you want some separation between this group and what's above it, don't apply \tcode{margin-top} to the whole group container.  
Instead apply it separately to each item inside the unit, 
so if some items wrap to the next line they will still have the appropriate top margin.
It's best to identify these types of issues early on.
\end{itemize}
\endgroup

\section{The Future of Web Components}
Polymer\index{Polymer} and Web Components\index{Web Components} are extremely powerful tools.
They are relatively immature as a technology but quickly improving.
Those who wish to deploy a large, full-featured web application based on these technologies may face some bumps in the road, 
at least until the standard is finalized and browser support is improved.
A project version below 1.0 is a general indicator of how ready its authors feel it is for production deployment.
Polymer's\index{Polymer} current version is at 0.5, soon to be 0.8, 
which suggests that 1.0 is not that far off.
Even the massively popular Angular\index{Angular} framework has announced that the controversial 2.0 re-write will be based on the Web Components standard,
so it appears that Web Components are here to stay~\cite{santiagoesteva2015}.

The Web Components\index{Web Components} initiative is much larger than just Polymer.
The draft is due to be formally adopted as an official HTML5\index{HTML!HTML5} standard by the World Wide Web Consortium\index{W3C}.
Although it will be a long time before full-speed native Web Component support exists in a majority of browsers worldwide, 
polyfill\index{polyfill} libraries can help bring these capabilities to current browsers.
However, the prospect of forcing visitors to download a several hundred KB library just to use basic DOM functionality is not very appealing to most web engineers from a performance and user experience\index{user experience (UX)} perspective.
The lack of native browser support may be a limiting factor in Web Component adoption in the near term.

In essence, Web Component's\index{Web Components} key problem at the moment is that one has to download a large JavaScript\index{JavaScript} framework\index{framework} in order to use it, 
yet one of its main premises is that it frees you from needing a JavaScript framework.
Of course, the path out of this chicken-and-egg problem is for browser vendors to fully implement native Web Components in their products,
a process which won't be complete until well after the standard is finalized by the W3C.
Given the enormous size of the job that the polyfill library must accomplish, the results obtained so far are impressive. 
Speakur\index{Speakur} has demonstrated that it's possible to create focused, extensible, yet full-featured Web Components right now with the help of tools like Polymer\index{Polymer} and Firebase\index{Firebase}.
